---
title: "TCGA Enrichment Analysis"

output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("STRINGdb")
library("knitr")
library("gplots")
library("RColorBrewer")
library("ggplot2")
```
## Load data
```{r loadData}
pathway <- readRDS(file = "use_case_data/tcga_data/pathway_comparison/selected_pathway_z2.rds")
pathway <- pathway$`K-3-L1-250`@nodes$node
z_score_matrix <- readRDS("use_case_data/tcga_data/data/string_z_score_counts.rds")
string_db <- STRINGdb$new(
  version = "11",
  species = 9606,
  input_directory = "stringdb800/",
  score_threshold = 800
)
```

## Visualize pathway
```{r visualize_pathway,fig.height = 6}
pathway_matrix <- z_score_matrix[z_score_matrix$STRING_id %in% pathway, ]
row.names(pathway_matrix) <- NULL

pathway_matrix <- tibble::column_to_rownames(pathway_matrix, var = "STRING_id")
pathway <- data.frame(STRING_id = row.names(pathway_matrix), z_score_mean = rowMeans(pathway_matrix[-1]))
pathway["color"] <- c("")
# FFFFFFFF -1 > Z- score < 2, C7FFD5 Z-Score > 2, 00FF00 Z-Score > 3
pathway$color[pathway$z_score_mean < -1] <- "#023e8a"
pathway$color[pathway$z_score_mean >= -1 & pathway$z_score_mean <= 3] <- "#FFFFFFFF"
pathway$color[pathway$z_score_mean > 3] <- "#fb8500"
# Upload payload information to STRING server
payload_id <- string_db$post_payload(pathway$STRING_id, colors = pathway$color)
#
# pdf(file = "use_case_data/tcga_data/plots/z_score_2/K-3-L1-250.pdf", width = 6, height = 7)
# string_db$plot_network(pathway$STRING_id,payload_id = payload_id)
# dev.off()
string_db$plot_network(pathway$STRING_id, payload_id = payload_id)

pathway <- string_db$add_proteins_description(pathway)
```

## Mean Z-score distribution of nodes in the pathway
```{r mean_z_score}
# Color palette for heatmap
pathway$preferred_name <- factor(pathway$preferred_name, level = pathway$preferred_name[order(pathway$z_score_mean)])
ggplot(pathway, aes(x = preferred_name, y = z_score_mean)) +
  geom_point() +
  geom_hline(yintercept = 1, linetype = "dashed", color = "#FFA400", size = 0.5) +
  geom_hline(yintercept = -1, linetype = "dashed", color = "#FFA400", size = 0.5) +
  geom_hline(yintercept = 2, linetype = "dashed", color = "#2A2A72", size = 0.7) +
  geom_hline(yintercept = -2, linetype = "dashed", color = "#2A2A72", size = 0.7) +
  geom_hline(yintercept = 4, linetype = "dashed", color = "purple", size = 1) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  labs(title = "Mean Z-score", y = "Z-score", x = "Gene")
```

## Compute Enrichment
The enrichment is computed in Gene Ontology (Process, Function and Component), KEGG and Reactome pathways, PubMed publications, UniProt Keywords, and PFAM/INTERPRO/SMART using an hypergeometric test and the FDR is calculated using Benjamini-Hochberg (Provided by the   stringDB package).
```{r compute_enrichmet}
# Set background
background <- z_score_matrix$STRING_id
string_db$set_background(z_score_matrix$STRING_id)
# Compute enrichment
enrichment <- string_db$get_enrichment(pathway$STRING_id)
write.table(enrichment, file='use_case_data/tcga_data/enrichment/enrichmment_table.tsv', row.names = F, quote=FALSE, sep='\t')
kable(enrichment)
```
## Visualize enriched terms 
```{r visualize_enrichment terms}
enrichment["gene_ratio"] <- enrichment$number_of_genes / nrow(pathway)

# GO biological process ####
bp <- enrichment[enrichment$category == "Process", ][1:20, ]
bp$description <- factor(bp$description, level = bp$description[order(bp$gene_ratio)])
bp_plot <- ggplot(bp, aes(y = description, x = gene_ratio, color = fdr)) +
  geom_point(aes(size = number_of_genes)) +
  scale_color_gradient(low = "#0353a4", high = "#ff8500") +
  labs(color = "FDR", size = "Gene count", x = "Percentage from all genes in the pathway", y = "Biological process") +
  ggtitle("Top 20 GO terms (BP)")+
  theme_minimal()

ggsave(bp_plot, file = "plots_and_figures/tcga/bp_enrichment.png")

#
bp <- enrichment[enrichment$category == "KEGG", ]
bp$description <- factor(bp$description, level = bp$description[order(bp$gene_ratio)])
bp_plot <- ggplot(bp, aes(y = description, x = gene_ratio, color = fdr)) +
  geom_point(aes(size = number_of_genes)) +
  scale_color_gradient(low = "#0353a4", high = "#ff8500") +
  labs(color = "FDR", size = "Gene count", x = "Percentage from all genes in the pathway", y = "Biological process") +
  ggtitle("Top 20 GO terms (BP)")+
  theme_minimal()




```







